### 2025-09-20 YEAR2/AER210_LectureNotes/main.tex
sept20

% set section counter from 15 
\setcounter{section}{14}
\paragraph{Note} the section numbering is based on Stewart's book.
\section{Double and Triple Integrals}
\begin{definition}[Double Integral]
    Let $f(x,y)$ be a function defined on a closed and bounded region $R$ in the $xy$-plane. The double integral of $f$ over $R$ is denoted by
    \begin{equation}
        \iint_R f(x,y) \, dA = \iint_R f(x,y) \, dA
    \end{equation}
    where $dA$ represents an infinitesimal area element in the region $R$. The double integral can be interpreted as the volume under the surface defined by $z = f(x,y)$ over the region $R$.
\end{definition}

@@

    Then, we can express the base region $R$ in polar coordinates:
    $$        
    x^2 + y^2 = r^2 \implies 0 \le r \le a, \quad 0 \le \theta \le 2\pi
    $$
    So we can set up:
    \begin{align*}
        I_z &= \rho_0 h \int_0^{2\pi} \int_0^a r^2 \cdot r \, dr \, d\theta \\
        &= \rho_0 h \int_0^{2\pi} \left[ \frac{r^4}{4} \right]_{r=0}^{r=a} \, d\theta = \frac{\pi \rho_0 h a^4}{2}
    \end{align*}
    This also leads to the well-known formula for the moment of inertia of a solid cylinder about its central axis:
    $$
        I_z = \frac{1}{2} M a^2
    $$
    where $M = \pi a^2 h \rho_0$ is the mass of the cylinder. 
\end{example}
\subsection{Triple Integrals in Cylindrical Coordinates}
\begin{definition}[Cylindrical Coordinates]
    The cylindrical coordinates of a point $P$ in 3-D space are given by the ordered triple $(r, \theta, z)$, where:
    \begin{itemize}
        \item $r$ is the distance from the $z$-axis to the projection of $P$ onto the $xy$-plane,
        \item $\theta$ is the angle between the positive $x$-axis and the line segment from the origin to the projection of $P$ onto the $xy$-plane,
        \item $z$ is the same as in rectangular coordinates, representing the height of point $P$ above the $xy$-plane.
    \end{itemize}
    The relationships between cylindrical coordinates and rectangular coordinates are given by:
    \begin{equation}
        x = r \cos \theta, \quad y = r \sin \theta, \quad z = z
    \end{equation}
    and conversely:
    \begin{equation}
        r = \sqrt{x^2 + y^2}, \quad \theta = \tan^{-1}\left(\frac{y}{x}\right), \quad z = z
    \end{equation}
\end{definition}

\begin{example}

\end{example}
\subsection{Triple Integrals in Spherical Coordinates}
\subsection{Change of Variables in Multiple Integrals} \label{sec:change_of_variables}
\chapter{Fluid Mechanics}
\end{document}

>>>>

% set section counter from 15 
\setcounter{section}{13}
\paragraph{Note} the section numbering is based on Stewart's book.

\section{Partial Derivatives}
\subsection{Taylor Series for Multivariable Functions}
\paragraph{Continuing from Calc II} We consider Taylor series of multivariable functions, and we would need to consider partial derivatives.

\begin{shaded}
    \subsubsection{Review. Taylor Series for Single Variable Functions}
    Let $f(x)$ be a function that is infinitely differentiable at $x = a$. The Taylor series of $f(x)$ about the point $x = a$ is given by:
    \begin{equation}
        f(x) = f(a) + f'(a)(x-a) + \frac{f''(a)}{2!}(x-a)^2 + \frac{f'''(a)}{3!}(x-a)^3 + \cdots = \sum_{n=0}^{\infty} \frac{f^{(n)}(a)}{n!} (x-a)^n
    \end{equation}

    Also, consider the taylor approximation of $f(x + \Delta x)$ about $x$: 
    \begin{equation}
        f(x + \Delta x) = f(x) + f'(x)\Delta x + \frac{f''(x)}{2!}(\Delta x)^2 + \frac{f'''(x)}{3!}(\Delta x)^3 + \cdots = \sum_{n=0}^{\infty} \frac{f^{(n)}(x)}{n!} (\Delta x)^n
    \end{equation}
\end{shaded}
\begin{definition}[Taylor Series for Multivariable Functions]
    Let $f(x,y)$ be a function that is infinitely differentiable at the point $(a,b)$. Consider the increment $\Delta x$ in the $x$-direction and $\Delta y$ in the $y$-direction centred at $(x_0,y_0)$. We have the following parametric equations:
    $$
        \begin{cases}
            x = x_0 + \Delta x t \\
            y = y_0 + \Delta y t
        \end{cases} \quad t \in [0,1]
    $$
    Define a new function $g(t) = f(x_0 + \Delta x t, y_0 + \Delta y t)$, which is a single-variable function in terms of $t$. We can then apply the Taylor series for single-variable functions to $g(t)$ about $t = 0$:
    $$
        g(t) = g(0) + g'(0)t + \frac{g''(0)}{2!}t^2 + \frac{g'''(0)}{3!}t^3 + \cdots = \sum_{n=0}^{\infty} \frac{g^{(n)}(0)}{n!} t^n
    $$
    Using the chain rule, we can compute the taylor series of $f(x,y)$ about the point $(x_0,y_0)$:
    \begin{align}
        f(x_0 + \Delta x, y_0 + \Delta y) &= f(x_0,y_0) + \left( \frac{\partial f}{\partial x} \Delta x + \frac{\partial f}{\partial y} \Delta y \right) + \nonumber \\
        &\quad \frac{1}{2!} \left( \frac{\partial^2 f}{\partial x^2} (\Delta x)^2 + 2\frac{\partial^2 f}{\partial x \partial y} \Delta x \Delta y + \frac{\partial^2 f}{\partial y^2} (\Delta y)^2 \right) + \cdots \\
        &= f(x_0,y_0) + \sum_{n=1}^{\infty} \sum_{k=0}^n \frac{1}{k!(n-k)!} \frac{\partial^n f}{\partial x^k \partial y^{n-k}} (\Delta x)^k (\Delta y)^{n-k}
    \end{align}
    Or equivalently, we can write:
    \begin{equation}
        f(x,y) = f(x_0,y_0) + \sum_{n=1}^{\infty} \sum_{k=0}^n \frac{1}{k!(n-k)!} \frac{\partial^n f}{\partial x^k \partial y^{n-k}} (x - x_0)^k (y - y_0)^{n-k}
    \end{equation}
    We could also use the fact:
    $$
        \frac{1}{k!(n-k)!} = \frac{1}{n!} \binom{n}{k}
    $$
    we rewrite the above equations, such that we can use the pascal's triangle to help us remember the coefficients:
    \begin{equation}
        f(x,y) = f(x_0,y_0) + \sum_{n=1}^{\infty} \frac{1}{n!} \sum_{k=0}^n \binom{n}{k} \frac{\partial^n f}{\partial x^k \partial y^{n-k}} (x - x_0)^k (y - y_0)^{n-k}
    \end{equation}

\end{definition}
\section{Multiple Integrals}
\begin{definition}[Double Integral]
    Let $f(x,y)$ be a function defined on a closed and bounded region $R$ in the $xy$-plane. The double integral of $f$ over $R$ is denoted by
    \begin{equation}
        \iint_R f(x,y) \, dA = \iint_R f(x,y) \, dA
    \end{equation}
    where $dA$ represents an infinitesimal area element in the region $R$. The double integral can be interpreted as the volume under the surface defined by $z = f(x,y)$ over the region $R$.
\end{definition}

@@

    Then, we can express the base region $R$ in polar coordinates:
    $$        
    x^2 + y^2 = r^2 \implies 0 \le r \le a, \quad 0 \le \theta \le 2\pi
    $$
    So we can set up:
    \begin{align*}
        I_z &= \rho_0 h \int_0^{2\pi} \int_0^a r^2 \cdot r \, dr \, d\theta \\
        &= \rho_0 h \int_0^{2\pi} \left[ \frac{r^4}{4} \right]_{r=0}^{r=a} \, d\theta = \frac{\pi \rho_0 h a^4}{2}
    \end{align*}
    This also leads to the well-known formula for the moment of inertia of a solid cylinder about its central axis:
    $$
        I_z = \frac{1}{2} M a^2
    $$
    where $M = \pi a^2 h \rho_0$ is the mass of the cylinder. 
\end{example}
\subsection{Triple Integrals in Cylindrical Coordinates}
\begin{definition}[Cylindrical Coordinates]
    The cylindrical coordinates of a point $P$ in 3-D space are given by the ordered triple $(r, \theta, z)$, where:
    \begin{itemize}
        \item $r$ is the distance from the $z$-axis to the projection of $P$ onto the $xy$-plane,
        \item $\theta$ is the angle between the positive $x$-axis and the line segment from the origin to the projection of $P$ onto the $xy$-plane,
        \item $z$ is the same as in rectangular coordinates, representing the height of point $P$ above the $xy$-plane.
    \end{itemize}
    The relationships between cylindrical coordinates and rectangular coordinates are given by:
    \begin{equation}
        x = r \cos \theta, \quad y = r \sin \theta, \quad z = z
    \end{equation}
    and conversely:
    \begin{equation}
        r = \sqrt{x^2 + y^2}, \quad \theta = \tan^{-1}\left(\frac{y}{x}\right), \quad z = z
    \end{equation}

    For the change of variables in triple integrals, the volume element $dV$ in cylindrical coordinates is given by:
    \begin{equation}
        dV = dA \, dz = r \, dr \, d\theta \, dz
    \end{equation}
\end{definition}

\begin{example}
    Consider the triple integral in the cononical reigion $V$ bounded by the cone $z = \sqrt{x^2 + y^2}$ and the plane $z=2$. Changing to cylindrical coordinates, we consider the projection of the cone onto the $xy$-plane as our polar base region $R$. Then, we consider how the cone change in the $z$ direction over $R$:
    $$
        z = r \implies r \le z \le 2
    $$
    So the region $V_{\text{polar}}$ is:
    $$
    V_{\text{polar}} = \{(r, \theta, z) \mid 0 \le r \le 2, 0 \le \theta \le 2\pi, r \le z \le 2\}
    $$
    So we can set up:
    $$
        \iiint_V f(x,y,z) \, dV = \int_0^{2\pi} \int_0^2 \int_r^2 f(r \cos \theta, r \sin \theta, z) \, r \, dz \, dr \, d\theta
    $$

    Let $f(x,y,z) = x^2 +. y^2 = r^2$. Then, we have:
    \begin{align*}
        \iiint_V (x^2 + y^2) \, dV &= \int_0^{2\pi} \int_0^2 \int_r^2 r^2 \cdot r \, dz \, dr \, d\theta \\
        &= \int_0^{2\pi} \int_0^2 r^3 (2 - r) \, dr \, d\theta = \frac{16\pi}{5}
    \end{align*}

\end{example}
\subsection{Triple Integrals in Spherical Coordinates}
\begin{definition}[Spherical Coordinates]
    The spherical coordinates of a point $P$ in 3-D space are given by the ordered triple $(\rho, \phi, \theta)$, where:
    \begin{itemize}
        \item $\rho$ is the distance from the origin to the point $P$,
        \item $\phi$ is the angle between the positive $z$-axis and the line segment from the origin to point $P$ (also known as the polar angle or colatitude),
        \item $\theta$ is the angle between the positive $x$-axis and the projection of the line segment from the origin to point $P$ onto the $xy$-plane (also known as the azimuthal angle).
    \end{itemize}
    The relationships between spherical coordinates and rectangular coordinates are given by:
    \begin{equation}
        x = \rho \sin \phi \cos \theta, \quad y = \rho \sin \phi \sin \theta, \quad z = \rho \cos \phi
    \end{equation}
    and conversely:
    \begin{equation}
        \rho = \sqrt{x^2 + y^2 + z^2}, \quad \phi = \cos^{-1}\left(\frac{z}{\sqrt{x^2 + y^2 + z^2}}\right), \quad \theta = \tan^{-1}\left(\frac{y}{x}\right)
    \end{equation}

    For the change of variables in triple integrals, the volume element $dV$ in spherical coordinates is given by:
    \begin{equation}
        dV = \rho^2 \sin \phi \, d\rho \, d\phi \, d\theta
    \end{equation}
\end{definition}

\begin{example}
    \textit{find the mass of a half sphere of radius $a$ that has a density $ k(2a - \rho)$, where $k$ is a constant and $\rho$ is the distance from the origin.}

    Let $\lambda = k(2a - \rho)$ be the density function. Then, we consider the half sphere $S$ in spherical coordinates:
    $$       
    S = \{(\rho, \phi, \theta) \mid 0 \le \rho \le a, 0 \le \phi \le \frac{\pi}{2}, 0 \le \theta \le 2\pi\}
    $$
    So we can set up:
    \begin{align*}
        M &= \iiint_S \lambda \, dV = \iiint_S k(2a - \rho) \, dV \\
        &= k \int_0^{2\pi} \int_0^{\frac{\pi}{2}} \int_0^a (2a - \rho) \rho^2 \sin \phi \, d\rho \, d\phi \, d\theta \\
        &= k \int_0^{2\pi} \int_0^{\frac{\pi}{2}} \left[ 2a \cdot \frac{\rho^3}{3} - \frac{\rho^4}{4} \right]_{\rho=0}^{\rho=a} \sin \phi \, d\phi \, d\theta \\
        &= k \int_0^{2\pi} \int_0^{\frac{\pi}{2}} \left( \frac{2a^4}{3} - \frac{a^4}{4} \right) \sin \phi \, d\phi \, d\theta \\
        &= k \cdot \frac{5a^4}{12} \int_0^{2\pi} \left[ -\cos \phi \right]_{\phi=0}^{\phi=\frac{\pi}{2}} \, d\theta = k \cdot \frac{5a^4}{12} \cdot 2\pi \\ 
        &= \frac{5\pi k a^4}{6}
    \end{align*}

\subsection{}
\subsection{Change of Variables in Multiple Integrals} \label{sec:change_of_variables}
\chapter{Fluid Mechanics}
\end{document}

### 2025-09-20 YEAR2/CHE260_LectureNotes/main.tex
sept20

\begin{example}[Pump at Constant Temperature]
    Consider a pump (a up pumping pump) that has 20$^\circ$C and 100 kPa at the inlet, and 20$^\circ$C and 600 kPa at the outlet. Assuming incompressible liquid water, and $v = 0.001\, \text{m}^3/\text{kg}$. A 10kg flow rate would have:
    $$
        w_{shaft} = h_2 - h_1 + g(z_2 - z_1)
    $$
    We can compute the enthalpy change:
    $$
        \Delta h = v (P_2 - P_1) 
    $$
    So the power with considerion of the change in potential energy is:
    $$
        W = m \Delta h + mg(z_2 - z_1) = 10 \times 0.001 \times (600 - 100) + 10 \times 9.81 \times (10 - 0) = 98.5\, \text{kW}
    $$
\end{example}
\chapter{Heat Transfer}

>>>>

\begin{example}[Pump at Constant Temperature]
    Consider a pump (a up pumping pump) that has 20$^\circ$C and 100 kPa at the inlet, and 20$^\circ$C and 600 kPa at the outlet. Assuming incompressible liquid water, and $v = 0.001\, \text{m}^3/\text{kg}$. A 10kg flow rate would have:
    $$
        w_{shaft} = h_2 - h_1 + g(z_2 - z_1)
    $$
    We can compute the enthalpy change:
    $$
        \Delta h = v (P_2 - P_1) 
    $$
    So the power with considerion of the change in potential energy is:
    $$
        W = m \Delta h + mg(z_2 - z_1) = 10 \times 0.001 \times (600 - 100) + 10 \times 9.81 \times (10 - 0) = 98.5\, \text{kW}
    $$
\end{example}

\subsection{Second Law of Thermodynamics and Entropy}
\paragraph{Motivation} We want to understand the direction of energy transfer and the efficiency of energy conversion processes. In an engine or a driven process, we want a measurement for the \textit{quality} of energy, and how much of it is \textit{usable}. This is important for understanding how to design efficient systems, and how to analyze thermodynamic processes.
\begin{definition}[Entropy]
    Entropy is an extensive property that changes when heat is transferred to or from a system. The entropy change is defined as:
    \begin{equation}
        dS = \frac{\delta Q_{rev}}{T}
    \end{equation}
    where $dS$ is the change in entropy, $\delta Q_{rev}$ is the reversible heat transfer, and $T$ is the absolute temperature. 
    Now for the total entropy change of a system, we have:
    \begin{equation}
        \Delta S_{system} = S_2 - S_1 = \int_{1}^{2} \frac{\delta Q_{rev}}{T}
    \end{equation}
    where $S_1$ and $S_2$ are the initial and final entropies of the system, respectively.
\end{definition}

\paragraph{Inevitable Entropy Generation} Consider a two systems undergoing heat transfer, we can model the entropy of the entire system as:
$$
    \Delta S_{universe} = \frac{Q_{rev}}{T_{hot}} - \frac{Q_{rev}}{T_{cold}}
$$
Since $T_{hot} > T_{cold}$, we have $\Delta S_{universe} > 0$. This means that for any process that involves heat transfer, the entropy of the universe always increases. This is a statement of the second law of thermodynamics.

\begin{theorem}[Second Law of Thermodynamics]
    The entropy in an isolated system always increases for any real process. This can be expressed as:
    \begin{equation}
        \Delta S > 0
    \end{equation}

    Entropy would generate until it reaches equilibrium, where $\Delta S_{universe} = 0$.
\end{theorem}
\chapter{Heat Transfer}

### 2025-09-20 YEAR2/ECE253_LectureNotes/main.tex
sept20

>>>>

\subsection{Decoders and Encoders}
\begin{definition}[Decoder]
    A decoder is a combinational logic circuit that converts binary information from $n$ input lines to a maximum of $2^n$ unique output lines. Each output line corresponds to one of the possible combinations of the input lines. When a specific combination of input lines is activated (set to 1), the corresponding output line is activated (set to 1), while all other output lines remain inactive (set to 0). Decoders are commonly used in applications such as memory address decoding, data multiplexing, and digital display systems.

\end{definition}
\begin{example}[2-bit 7-Segment Display]
    A 7-segment display is an electronic display device that consists of seven individual segments (labeled $h_0$ to $h_6$) that can be illuminated in different combinations to represent numerical digits (0-9) and some alphabetic characters. The truth table for binary inputs $x_0, x_1, x_2, x_3$ (only $x_0$ and $x_1$ are used in this demonstration) to control the 7-segment display is shown below (note that 0 represents an active segment):
    \begin{table}[h!]
        \centering
        \begin{tabular}{|c|c||c|c|c|c|c|c|c|}
            \hline
            $x_1$ & $x_0$ & $h_0$ & $h_1$ & $h_2$ & $h_3$ & $h_4$ & $h_5$ & $h_6$ \\
            \hline
            0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 \\ % 0
            0 & 1 & 1 & 0 & 0 & 1 & 1 & 1 & 1 \\ % 1
            1 & 0 & 0 & 0 & 1 & 0 & 0 & 1 & 0 \\ % 2
            1 & 1 & 0 & 0 & 0 & 0 & 1 & 1 & 0 \\ % 3
        \hline
        \end{tabular}
        \caption{Truth Table for 7-Segment Display (0-3)}
        \label{tab:7_segment_display_truth_table}
    \end{table}
    So we have:
    \begin{align*}
        h_0 &= \overline{x_1}{x_0} \\
        h_1 &= 0 \\
        h_2 &= x_1\overline{x_0} \\
        h_3 &= h_0 = \overline{x_1}{x_0} \\
        h_4 &= x_0 \\
        h_5 &= x_0 + x_1 \\
        h_6 &= \overline{x_1} \\
    \end{align*}
    The Verilog implementation is as follows:
    \begin{verbatim}
    module seven_segment(input logic [1:0] x,
                         output logic [6:0] h);
        assign h[0] = ~x[1] & x[0];
        assign h[1] = 1'b0;
        assign h[2] = x[1] & ~x[0];
        assign h[3] = ~x[1] & x[0];
        assign h[4] = x[0];
        assign h[5] = x[0] | x[1];
        assign h[6] = ~x[1];
    endmodule
    \end{verbatim}

\begin{definition}[If-Else Statement]
    An if-else statement is a conditional statement that allows for branching in the execution of code based on the evaluation of a boolean expression. In Verilog, if-else statements are used to model combinational logic and can be used within always blocks or initial blocks. The syntax for an implementation of mux using an if-else statement in Verilog is as follows:
    \begin{verbatim}
        module mux(input logic x, y, s,
                   output logic f);
            always_comb // Always @(x or y or s) is equivalent.
            // The sensitivity list gives the signals
            // that can trigger the always block.
            begin
                if (s == 1'b0) // If statements must go inside always block.
                    f = x; // Keyword 'assign' not used in always block.
                else
                    f = y;
                end
            end
        endmodule
    \end{verbatim}
    \textbf{Analogy} The always block is like the \verb|useEffect(() => { ... }| hook in React. The sensitivity list is like the dependency array in React.
\end{definition}

\begin{definition}[Case Statement]
    A case statement is a control flow statement that allows for multi-way branching based on the value of a single expression. In Verilog, case statements are used to model combinational logic and can be used within always blocks. The syntax for an implementation of a full 7-segment display using a case statement in Verilog is as follows:
    \begin{verbatim}
        module seven_segment_case(input logic [3:0] x,
                                  output logic [6:0] h);
            always_comb begin
                case (x) 
                    4'b0000: h = 7'b1000000; // 0
                    4'b0001: h = 7'b1111001; // 1
                    4'b0010: h = 7'b0100100; // 2
                    4'b0011: h = 7'b0110000; // 3
                    4'b0100: h = 7'b0011001; // 4
                    4'b0101: h = 7'b0010010; // 5
                    4'b0110: h = 7'b0000010; // 6
                    4'b0111: h = 7'b1111000; // 7
                    4'b1000: h = 7'b0000000; // 8
                    4'b1001: h = 7'b0010000; // 9
                    default: h = 7'b1111111; // All segments off for invalid input
                endcase
            end
        endmodule
    \end{verbatim}

### 2025-09-20 YEAR2/MAT292_LectureNotes/main.tex
sept20

                \node at (axis cs:2,5) [anchor=north east] {\( x = \frac{c}{d} \)};
                % Fixed point marker
                \node at (axis cs:2,3) [circle,fill,inner sep=1.5pt,label=above right:{\(\left(\frac{c}{d}, \frac{a}{b}\right)\)}] {};
                \end{axis}
            \end{tikzpicture}
            \caption{Nullclines of the Lotka-Volterra model showing the prey nullcline (horizontal line) and predator nullcline (vertical line). The intersection point represents the coexistence equilibrium.}
        \end{figure}

@@

    \end{itemize}
\end{example}

>>>>

                \node at (axis cs:2,5) [anchor=north east] {\( x = \frac{c}{d} \)};
                % Fixed point marker
                \node at (axis cs:2,3) [circle,fill,inner sep=1.5pt,label=above right:{\(\left(\frac{c}{d}, \frac{a}{b}\right)\)}] {};
                \end{axis}
            \end{tikzpicture}
            \caption{Nullclines of the Lotka-Volterra model showing the prey nullcline (horizontal line) and predator nullcline (vertical line). The intersection point represents the coexistence equilibrium.}
        \end{figure}
        And a figure for the vector field is shown below:
        \begin{figure}[h!]
            \centering
            \includegraphics[width=0.6\linewidth]{predprey.png}
            \caption{Vector field of the Lotka-Volterra model showing the circular flow around
            the fixed point \(\left(\frac{c}{d}, \frac{a}{b}\right)\).}
        \end{figure}
\end{example}

@@

    \end{itemize}

    A vector field of the system is shown below:
        \begin{figure}[h!]
            \centering
            \includegraphics[width=0.6\linewidth]{pendulum.png}
            \caption{Vector field of the simple pendulum system showing the fixed points at \( (n\pi, 0) \).}
        \end{figure}

\paragraph{Systems of ODEs as Matrices}
A system of ODEs can be expressed in matrix form as:
$$\mathbf{x}' = A\mathbf{x} + \mathbf{b}(t)
$$where \( \mathbf{x} \) is the state vector, \( A \) is a matrix of coefficients, and \( \mathbf{b}(t) \) is a vector of functions.

\begin{definition}[General System of ODEs]
    WLOG, for a system of 2 ODEs with $x$ and $y$, we write:
    $$\begin{cases}
    x' = f(t, x, y) \\
    y' = g(t, x, y)
    \end{cases}$$
    where \( f \) and \( g \) are given functions:
    $$
    f(t, x, y) = p_{11}(t)x + p_{12}(t)y + g_1(t) \\
    g(t, x, y) = p_{21}(t)x + p_{22}(t)y + g_2(t)
    $$
    We can express this in matrix form as:
    \begin{equation}
    \begin{bmatrix}x' \\ y'
    \end{bmatrix} = \begin{bmatrix}
    p_{11}(t) & p_{12}(t) \\
    p_{21}(t) & p_{22}(t)
    \end{bmatrix} \begin{bmatrix}x \\ y
    \end{bmatrix} + \begin{bmatrix}g_1(t) \\ g_2(t)
    \end{bmatrix}
    \end{equation}

    Furthermore, we generalize this to \( n \) dimensions, and the inhomogeneous, non-autonomous system of linears ODEs is given by:
    \begin{align*}
    \mathbf{x}'_1 &= \sum^n_{j=1} P_{1j}(t)x_j + g_1(t) \quad i = 1, 2, \ldots, n \\
     & \vdots \\
    \mathbf{x}'_n &= \sum^n_{j=1} P_{nj}(t)x_j + g_n(t)
    \end{align*}
    or in matrix form:
    \begin{equation}
    \mathbf{x}' = A(t)\mathbf{x} + \mathbf{G}(t)
    \end{equation}
    where \( A(t) \) is an \( n \times n \) matrix of functions, \( \mathbf{x} \) is the state vector, and \( \mathbf{G}(t) \) is the inhomogeneous term.
\end{definition}

\begin{theorem}[Existence and Uniqueness for Systems of ODEs]
    If $p_{ij}(t)$ and $g_i(t)$ are continuous on an open interval $I = (a, b)$ containing $t_0$ for all $i, j = 1, 2, \ldots, n$, then the initial value problem
    $$\mathbf{x}' = A(t)\mathbf{x} + \mathbf{G}(t), \quad \mathbf{x}(t_0) = \mathbf{x}_0$$
    has a unique solution on the interval \( I \).
\end{theorem}

\begin{example}[Linear, homogeneous and autonomous]
    Consider the simple ODE system:
    $$
        \textbf{x}' = A\textbf{x} +\textbf{b}
    $$
    We cosidner the fix points by setting \( \textbf{x}' = 0 \):
    $$A\textbf{x} + \textbf{b} = 0 \implies A\textbf{x} = -\textbf{b}$$
    If \( A \) is invertible, then the fixed point is given by:
    $$\textbf{x}^* = -A^{-1}\textbf{b}$$
    If \( A \) is not invertible, then there are either no fixed points or infinitely many fixed points. Now, we let $\textbf{y} = \textbf{x} + A^{-1}\textbf{b}$, then we have:
    $$\textbf{y}' =\textbf{x}'$$
    we have:
    $$Ay = A\textbf{x} + \textbf{b} = \textbf{x}' = \textbf{y}'$$
    Thus, the system reduces to:
    $$\textbf{y}' = A\textbf{y}$$
    which is a linear, homogeneous, and autonomous system. The fixed point of this system is given by:
    $$\textbf{y}^* = 0$$
\end{example}

\begin{example}[Simple ODEs]
    Consider the system of ODEs:
    $$
    \textbf{x}' = \begin{bmatrix} 1 & 0 \\ 0 & -2
    \end{bmatrix}\textbf{x} 
    $$
    and the initial condition:
    $$
    \textbf{x}(0) = \begin{bmatrix} x_0 \\ y_0
    \end{bmatrix}
    $$
    Then we can consider the solutions:
    $$x_1(t) = x_0 e^{t}$$
    $$x_2(t) = y_0 e^{-2t}$$
    or we can write the solution in vector form:
    $$\textbf{x}(t) = x_0 \begin{bmatrix}
        1 \\ 0
    \end{bmatrix} e^{t} + y_0 \begin{bmatrix}
        0 \\ 1
    \end{bmatrix} e^{-2t}
    $$

    The fixed point and nullclines are given by:
    $$\textbf{x}^* = \begin{bmatrix}0 \\ 0
    \end{bmatrix}$$
    The nullclines are given by:
    \begin{itemize}
        \item \( x' = 0 \): \( x = 0 \) (vertical line)
        \item \( y' = 0 \): \( y = 0 \) (horizontal line)
    \end{itemize}

    \textbf{Stability of Nullclines} The $x$-nullcline since it goes to the origin as $t \to -\infty$ and away from the origin as $t \to \infty$, while the $y$-nullcline goes to the origin as $t \to \infty$ and away from the origin as $t \to -\infty$. 
        \begin{figure}[h!]
            \centering
            \begin{tikzpicture}
                \begin{axis}[
                    axis lines = middle,
                    view     = {0}{90},
                    xlabel = \( x \),
                    ylabel = \( y \),
                    ymin = -3, ymax = 3,
                    xmin = -3, xmax = 3,
                    domain=-3:3,
                    samples=200,
                    width=0.6\linewidth,
                    height=6cm,
                    grid=both,
                    minor tick num=1,
                    major grid style={line width=.2pt,draw=gray!50},
                    minor grid style={line width=.1pt,draw=gray!20},
                ]
                % x-nullcline (vertical line)
                \addplot[blue, thick] coordinates {(0,-3) (0,3)};
                % y-nullcline (horizontal line)
                \addplot[red, thick] coordinates {(-3,0) (3,0)};
                % Fixed point marker
                \node at (axis cs:0,0) [circle,fill,inner sep=1.5pt,label=above right:{\((0, 0)\)}] {};
                \end{axis}
            \end{tikzpicture}
            \caption{Phase portrait of the simple ODE system showing the nullclines and the fixed point at the origin.}
        \end{figure}
        \textbf{Note - Stability} The stability here really means that the trajectories approach or move away from the fixed point along the nullclines, this does not mean for perturbation along the nullclines.
        \textbf{Saddle Node} The fixed point is a saddle node (since one eigenvalue is positive and the other is negative). The trajectories in the phase plane are shown below:
\end{example}

\begin{example}[Linear Algebra]
    Consider the system of ODEs:
    $$
    \textbf{x}' = A\textbf{x}
    $$
    where \( A \) is a constant matrix. Consider:
    \begin{align*}
        \textbf{x} &= \vec{v} e^{\lambda t} \\
        \textbf{x}' &= \lambda \vec{v} e^{\lambda t} \\
        &= A\vec{v} e^{\lambda t} = A\textbf{x}
    \end{align*}
    Thus, we have:

    \begin{equation}
    A\vec{v} = \lambda \vec{v}
        \end{equation}
    which is the eigenvalue problem. 
\end{example}

\begin{example}[Superposition]

    Assume $x_1(t)$ and $x_2(t)$ are solutions to the system of ODEs:
    $$\textbf{x}' = A\textbf{x}$$
    Then, any linear combination of these solutions is also a solution:
    $$\textbf{x}(t) = c_1 x_1(t) + c_2 x_2(t)$$
    where \( c_1 \) and \( c_2 \) are constants. This is due to the linearity of the system.
\end{example}

\begin{definition}[General Eigenproblem for n-dim ODEs System]
    Combining the above ideas, consider the system:
    $$\textbf{x}' = A\textbf{x}$$
    where \( A \in \mathbb{R}^{n \times n} \). We define the \textit{specturm} as:
    \begin{equation}
    \sigma(A) = \{(\lambda_i, \vec{v}_i) \in \mathbb{C} \times \mathbb{C}^n : A\vec{v}_i = \lambda_i \vec{v}_i, i = 1, 2, \ldots, n\}
    \end{equation}
    where \( \lambda_i \) are the eigenvalues and \( \vec{v}_i \) are the corresponding eigenvectors. The general solution to the system can be expressed as:
    \begin{equation}
            \textbf{x}(t) = \sum_{i=1}^{n} c_i \vec{v}_i e^{\lambda_i t}
    \end{equation}
    where \( c_i \) are constants determined by the initial conditions.
\end{definition}

\begin{example}
    Consider the following system of ODEs:
    $$
    \textbf{x}' = \begin{bmatrix} 0 & 1 \\ 1 & 0
    \end{bmatrix}\textbf{x}
    $$
    We find the eigenvalues and eigenvectors by solving the characteristic polynomial:
    \begin{align*}
    \det(A - \lambda I) &= \det\begin{bmatrix} -\lambda & 1 \\ 1 & -\lambda
    \end{bmatrix} \\
    &= \lambda^2 - 1 = 0 \\
    \implies \lambda_1 &= 1, \quad \lambda_2 = -1
    \end{align*}
    The corresponding eigenvectors are:
    \begin{align*}
    \vec{v}_1 &= \begin{bmatrix} 1 \\ 1
    \end{bmatrix}, \quad \vec{v}_2 = \begin{bmatrix} -1 \\ 1
    \end{bmatrix}
    \end{align*}
    Thus, the general solution to the system is:
    $$\textbf{x}(t) = c_1 \begin{bmatrix} 1 \\ 1
    \end{bmatrix} e^{t} + c_2 \begin{bmatrix} -1 \\ 1
    \end{bmatrix} e^{-t}$$
    where \( c_1 \) and \( c_2 \) are constants determined by the initial conditions.

    Consider the phase portait. If we start on the span of $ [-1, 1]^T$, then we go to the origin as $t \to \infty$. If we start on the span of $ [1, 1]^T$, then we go away from the origin as $t \to \infty$. Then, the trivial fixed point is a saddle node.
\end{example}
